### 2021-08-21 doc/thesis/chapters/dsp.tex
hls done

\noindent
The instantaneous frequency $f_i$ of an FM signal can be described as
\begin{equation}
  f_i = f_c + \Delta f \cdot m(t)
\end{equation}
where $f_c$ is the carrier- or center frequency, $\Delta f$ is the maximum frequency deviation and $m(t)$ is the information or message signal that is to be transmitted.
Simply looking at this equation, the instantaneous frequency varies in the range between $f_c \pm \Delta f$, which is sometimes also called the swing.
Considering the relationship between frequency and angle, as described above, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
\begin{equation}
  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int_{0}^{t} m(\tau) d\tau \Big),
  \label{equ:fm_func}
\end{equation}
where $A_c$ is the amplitude of the resulting FM signal and $k_f$ is the frequency sensitivity factor.
This sensitivity factor has a direct relationship with the modulation index $\beta$.
\begin{equation}
  \beta = \frac{\Delta f}{f_m} = \frac{k_f A_m}{f_m},
\end{equation}
where $f_m$ is the highest existing frequency, or the bandwidth of the information signal.
% More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm
Equation~\eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
A widely used example application of FM is broadcast radio, where audio streams are transmitted.
An audio stream can be described as a cosine wave.
Therefore, the information signal that is to be transmitted in broadcast radio can be represented as a cosine wave in the form of
\begin{equation}
  m(t) = A_m\ cos(2 \pi f_m t).
\end{equation}
By inserting this message signal into the generic FM Equation~\eqref{equ:fm_func}, the final formula for an FM signal transforms into the following form.
\begin{equation}
  y(t) = A_c\ cos \Big(2 \pi f_c t + \beta\ sin(2 \pi f_m t)\Big )
\end{equation}
Several formulas, as well as their derivations are described according to \cite[pg.54-55]{MatinMohammadA2018CSfE}.

@@

The differentiated signal still consists of a FM part, as explained above, using Equation~\eqref{equ:fm_demod_discriminator}.
In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
A block diagram is shown in Figure~\ref{fig:bd_freq_discriminator}.\\

>>>>

\noindent
The instantaneous frequency $f_i$ of an FM signal can be described as
\begin{equation}
  f_i = f_c + \Delta f \cdot m(t)
\end{equation}
where $f_c$ is the carrier- or center frequency, $\Delta f$ is the maximum frequency deviation and $m(t)$ is the information or message signal that is to be transmitted.
Simply looking at this equation, the instantaneous frequency varies in the range between $f_c \pm \Delta f$, which is sometimes also called the swing.
Considering the relationship between frequency and angle, and after some simple substitutions which will not be described into detail here, the equation for a generic frequency modulated wave is the following
\begin{equation}
  s(t) = A_c\ cos \Big( 2 \pi f_c t + 2 \pi k_f \int_{0}^{t} m(\tau) d\tau \Big),
  \label{equ:fm_func}
\end{equation}
where $A_c$ is the amplitude of the resulting FM signal and $k_f$ is the frequency sensitivity factor.
This sensitivity factor has a direct relationship with the modulation index $\beta$.
\begin{equation}
  \beta = \frac{\Delta f}{f_m} = \frac{k_f A_m}{f_m},
\end{equation}
where $f_m$ is the highest existing frequency, or the bandwidth of the information signal.
% More info on modulation index: https://fas.org/man/dod-101/navy/docs/es310/FM.htm
Equation~\eqref{equ:fm_func} describes a frequency modulated signal with a generic message signal $m(t)$.
A widely used example application of FM is broadcast radio, where audio streams are transmitted.
An audio stream can be described as a cosine wave.
Therefore, the information signal that is to be transmitted in broadcast radio can be represented as a cosine wave in the form of
\begin{equation}
  m(t) = A_m\ cos(2 \pi f_m t).
\end{equation}
By inserting this message signal into the generic FM Equation~\eqref{equ:fm_func}, the final formula for an FM signal transforms into the following form.
\begin{equation}
  y(t) = A_c\ cos \Big(2 \pi f_c t + \beta\ sin(2 \pi f_m t)\Big )
\end{equation}
Several formulas, as well as their derivations are described according to \cite[pg.54-55]{MatinMohammadA2018CSfE}.

@@

The differentiated signal still consists of a FM part, as explained earlier, using Equation~\eqref{equ:fm_demod_discriminator}.
In order to remove this high-frequeny part and extract the low-frequency envelope, the signal needs to be fed into an envelope detector.
A block diagram is shown in Figure~\ref{fig:bd_freq_discriminator}.\\

### 2021-08-21 doc/thesis/chapters/hls.tex
hls done

The general significance of the simulation testbench for a HDL design is explained here, as well as the different levels of abstraction in which a HLS testbench can be executed.
These different levels of abstraction in the testbench execution are a major difference between a VHDL testbench and an HLS testbench.
The most important impact thereof is the execution speed -- the time it takes to simulate a number of test cases on the DUT.\\

Section~\ref{sec:hls:workflow} above already mentioned the topics of testbenches and simulation types.
The following descriptions shall be seen as a comprehensive extension around these topics.

@@

Thus, in order to allow debugging and verification of HDL code, the technology of simulation was invented.
Here, the HDL is also translated into a netlist that represents all the parallel processes just like in hardware, but the subsequent step of synthesis is not taken.
Therefore, simulation comes to a result much faster.
Since HDL designs actually only consist of electrical wires and some basic block primitives, like flipflops or lookup tables, it is difficult to verify correct values.
They would need to be compared in their binary bit-pattern format, at the correct point in time.
Of course the HDL language provides support with different datatypes, that abstract the binary number representation and make the numbers human-readable.
However, it remains an enormous effort to manually write testing code to verify a bus interface, for example.
In order to simplify such complex operations, various testbench frameworks and libraries are available to developers.

@@

In C Simulation, the HLS compiler is used to create an executable binary that includes the testbench code, as well as the IP design of the DUT.
This binary can be run just like any other program.
The most important note here is, that the HLS code is \textit{not} transformed into RTL at this point.
Instead, the HLS code is treated as a regular \cplusplus\ code, that uses some libraries, like the fixed-point library.
This means, that the HLS IP cores' top-level function is used as a regular function call in the testbench.\\

@@

\noindent
This type of simulation takes a significantly longer time to execute, because timing information now actually needs to be considered for the RTL code.

>>>>

The general significance of the simulation testbench for an HDL design is explained here, as well as the different levels of abstraction in which an HLS testbench can be executed.
These different levels of abstraction in the testbench execution are a major difference between a VHDL testbench and an HLS testbench.
The most important impact thereof is the execution speed -- the time it takes to simulate a number of test cases on the DUT.\\

Section~\ref{sec:hls:workflow} already mentioned the topics of testbenches and simulation types briefly, in the context of workflow.
The following descriptions are a comprehensive extension around these topics.

@@

Thus, in order to allow debugging and verification of HDL design code, the technology of simulation was invented.
Here, the time-consuming step of synthesis is not taken.
Instead, the HDL simulator software is able to simulate the code in a way, that can represent parallel processes, just like they exist in hardware.
Therefore, simulation comes to a result much faster.
Since HDL designs actually only consist of electrical wires and some basic block primitives, like flipflops or lookup tables, it is difficult to verify correct values.
They would need to be compared in their binary bit-pattern format, at the correct point in time.
Of course the HDL language provides support with different data types, that abstract the binary number representation and make the numbers human-readable.
However, it remains an enormous effort to manually write testing code to verify bus interfaces, for example, because they consist of a large number of signals.
In order to simplify such complex verification operations, various testbench frameworks and libraries are available to developers.

@@

In C Simulation, the HLS compiler is used to create an executable binary that includes the testbench code, as well as the IP design of the DUT.
This binary can be run just like any other program.
The most important note here is, that the HLS code is \textit{not} transformed into RTL at this point.
Instead, the HLS code is treated as a regular \cplusplus\ code, that uses some libraries, like the fixed point library.
This means, that the HLS IP core's top-level function is used as a regular function call in the testbench.\\

@@

\noindent
This type of simulation takes a significantly longer time to execute, because timing information now actually needs to be considered for the RTL code, which is done by using an HDL simulator.

### 2021-08-21 doc/thesis/chapters/implementation.tex
hls done

The Matlab model is therefore developed by keeping in mind this limitation.
Matlab provides special datatypes for fixed point number representation.
However, these were not used in this implementation.
Instead, everything is calculated in floating point representation, using the standard \textit{double} datatype.
Then, in certain positions in the DSP chain, the numbers are rounded to the resolution of a fixed point datatype.
A fixed point format of \textit{2.14}, meaning 2 bit for integer- and 14 bit for fractional number representation is chosen.
This is sufficient for the FM receiver and could potentially be lowered to a lower bitwidth.
Again, an optimization of the DSP chain is not the focus of this thesis.\\

>>>>

The Matlab model is therefore developed by keeping in mind this limitation.
Matlab provides special data types for fixed point number representation.
However, these were not used in this implementation.
Instead, everything is calculated in floating point representation, using the standard \textit{double} data type.
Then, in certain positions in the DSP chain, the numbers are rounded to the resolution of a fixed point data type.
A fixed point format of \textit{2.14}, meaning 2 bit for integer- and 14 bit for fractional number representation is chosen.
This is sufficient for the FM receiver and could potentially be lowered to a lower bitwidth.
Again, an optimization of the DSP chain is not the focus of this thesis.\\
